Analysis_Report

### 2019-04-11 analysis/analysis_report.rmd
up to omission trial reward analysis

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(R.matlab)
library(afex)
library(emmeans)
library(here)
library(patchwork)
library(cowplot)
library(BayesFactor)
```

>>>>

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(R.matlab)
library(afex)
library(emmeans)
library(here)
library(patchwork)
library(cowplot)
library(BayesFactor)
library(hrbrthemes)
library(ggsignif)

vmacCols <- vmac_pal()
my_theme <- theme_ipsum_rc(grid = F, axis_title_just = "c", axis_title_size = 14, axis_text_size = 12, strip_text_size = 14, plot_title_size = 14, subtitle_size = 14, ticks = T) +
  theme(legend.position = "none", 
        plot.margin = margin(.5,1,.5,1,"cm"),
        axis.line = element_line(colour = "black", size = .5),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_line(colour = "black", size = .5),
        #aspect.ratio = 1.1,
        plot.tag = element_text(size = 20, family = font_rc),
        axis.title = element_text(margin = margin(12,1,1,1, unit = "cm")),
        axis.text.x = element_text(margin = margin(t = 4, unit = 'pt'), colour = 'black'),
        axis.text.y = element_text(margin = margin(r = 4, unit = 'pt'), colour = 'black'),
        axis.title.x = element_text(margin = margin(t = 8, unit = 'pt')),
        axis.title.y = element_text(margin = margin(r = 8, unit = 'pt')),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14))

```

# Behavioural Analysis

```{r Load Behavioural Data, include=FALSE}
files <- dir(path = here::here("analysis", "data", "raw_data", "BehavData"), pattern = "*.mat")

filedata <- here::here("analysis", "data", "raw_data", "BehavData", files) %>%
  map(readMat) %>%
  map("DATA")
```

```{r Tidy Behavioural Data, include = FALSE}
#This is some messy code that lets us tidy the exptTrialInfo data from the nested .mat struct files

subNums <- filedata %>%
  map(1) %>%
  flatten() %>%
  map_dbl(1)

exptdata <- filedata %>%
  map(24) %>%
  map(2) %>%
  modify(as.data.frame)

counterbals <- filedata %>%
  map(2) %>%
  flatten() %>%
  map_dbl(1)

ages <- filedata %>%
  map(10) %>%
  flatten() %>%
  map_dbl(1)

ages_summ <- ages %>%
  tibble() %>% 
  rename("age" = ".")%>%
  summarise(m = mean(age, na.rm = T), stdev = sd(age))

bonuses <- filedata %>%
  map(17) %>%
  flatten() %>% 
  map_dbl(1)

bonuses_sum <- bonuses %>%
  tibble() %>% 
  rename("bonus" = ".") %>%
  summarise(m = mean(bonus, na.rm = T), stdev = sd(bonus))

genders <- filedata %>%
  map(11) %>%
  flatten()  %>%
  map_chr(1) %>%
  tolower()

genders_summ <- genders %>%
  as.factor() %>% 
  tibble() %>% 
  rename("gender" = ".") %>%
  count(gender)

exptdata <- Map(cbind, exptdata, sub = subNums, counterbal = counterbals, age = ages, gender = genders) %>%
  bind_rows()

colnames(exptdata)[1:24] <- c("block", "trial", "trialCounter", "trials_since_break", "targetLoc", "targetType", "distractLoc", "distractShape", "distractType", "fixationTime", "fixationPropGoodSamples", "fixationTimeout", "trialPropGoodSamples", "timeout", "softTimeoutTrial", "omissionTrial", "rt", "trialPay", "sessionPay", paste("timeOnLoc", c(1:5)))

exptdata <- exptdata %>%
  mutate(condition = if_else(counterbal %in% c(1,2), "Singleton Search", "Feature Search"))

exptdata <- exptdata %>% 
  mutate(gender = tolower(gender)) %>% 
  mutate(dType = factor(distractType, levels = c(1,2,3), labels = c("High", "Low", "Absent")))
```

```{r Remove participants with poor gaze data}
propCutoff <- .5
allTracks <- exptdata %>%
  group_by(sub) %>%
  summarise(fixationProp = mean(fixationPropGoodSamples),
            trialProp = mean(trialPropGoodSamples))

poortracks <- allTracks %>%
  filter(fixationProp < propCutoff | trialProp < propCutoff)

allTracks <- allTracks %>%
  filter(!sub %in% poortracks$sub)

allTracks_summ <- allTracks %>%
  summarise(mean_fixation_prop = round(mean(fixationProp)*100, 1), sd_fixation_prop = round(sd(fixationProp)*100, 1), mean_trial_prop = round(mean(trialProp)*100, 1), sd_trial_prop = round(sd(trialProp)*100,1))

exptdata <- exptdata %>%
 filter(!sub %in% poortracks$sub)
```

We removed data for any participants who did not have sufficient quality gaze data. If participants have an average of less than `r propCutoff*100`% valid gaze data during the fixation period or during a trial, they are removed from the data set.

`r length(poortracks)` participants were removed due to having a poor track.

```{r Remove participants for any other reasons}

remove_p_numbers <- c(37)

#  check to make sure that everyone has completed the full 480 trials
didNotComplete <- exptdata %>%
  group_by(sub) %>%
  filter(block > 0) %>%
  tally() %>%
  filter(n < 480) %>%
  select(sub)

# remove those that did not complete the full expt
remove_p_numbers <- c(remove_p_numbers, didNotComplete$sub)

exptdata <- exptdata %>%
  filter(!sub %in% remove_p_numbers)
```

`r length(remove_p_numbers)` participants were removed for other reasons (participant #37 they thought that they needed to look at the coloured circle in order to earn rewards, participant #26 and #85 did not complete the experiment due to difficulty tracking their eyes).

Then we remove the first two trials of the experiment, the first two trials after breaks, and timeouts from the data (in line with previous VMAC analysis protocols):

```{r Additional data exclusions}
# first we determine the percentage of trials that were timeouts (for reporting in the MS)
hardtimeouts <- exptdata %>%
  group_by(timeout) %>%
  summarise(n = n()) %>%
  mutate(perc = round(n/sum(n)*100, 1))

exptdata <- exptdata %>%
  filter(trial > 2 & trials_since_break > 2) %>% #removing first two trials and trials after breaks
  filter(timeout != 1) # removing timeouts
```

## Omission Trials
```{r Omission Trial Means}
exptdata_means_byparticipant <- exptdata %>%
  ungroup() %>%
  group_by(sub, condition, dType) %>%
  summarise(m = mean(omissionTrial))

exptdata_means_bycondition <- summarySEwithin2(data = exptdata_means_byparticipant, measurevar = "m", betweenvars = "condition", withinvars = "dType", idvar = "sub") %>%
  select(-ends_with("Normed"))
```

```{r Omission Trial Graph}

omissionPlot_singleton <- exptdata %>%
  filter(condition == "Singleton Search") %>%
  group_by(sub, dType) %>%
  summarise(m = mean(omissionTrial)) %>%
  ggplot(aes(x = dType, y = m, fill = dType, shape = dType)) +
  geom_errorbar(data = filter(exptdata_means_bycondition, condition == "Singleton Search"), aes(x = dType, y = m, ymin = m-se, ymax = m+se), colour = "black", width = .15) +
  geom_line(aes(group = sub), colour = "black", alpha = .1, linetype = "solid", size = .5) +
  geom_line(data = filter(exptdata_means_bycondition, condition == "Singleton Search"), aes(x = dType, y = m, group = 1), colour = "black") + 
  geom_point(data = filter(exptdata_means_bycondition, condition == "Singleton Search"), aes(x = dType, y = m), colour = "black", size = 3) +
  geom_signif(comparisons = list(c("High", "Low"), c("Low", "Absent")), annotations = c("***","*"), y_position = c(.61,.57), colour = "black", tip_length = 0) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), expand = c(0,0), limits = c(0,.8)) +
  scale_x_discrete(labels = c("High\nreward", "Low\nreward", "Distractor\nabsent")) +
  scale_fill_manual(values = vmacCols) +
  scale_colour_manual(values = vmacCols) +
  scale_shape_manual(name = "Trial Type", guide = guide_legend(), values = c(21,22,23)) +
  labs(y = "% Reward Omissions", x = "Trial Type", tag = "A", title = "Singleton Search") +
  my_theme

omissionPlot_feature <- 
  exptdata %>%
  filter(condition == "Feature Search") %>%
  group_by(sub, dType) %>%
  summarise(m = mean(omissionTrial)) %>%
  ggplot(aes(x = dType, y = m, fill = dType, shape = dType)) +
  geom_line(aes(group = sub), colour = "black", alpha = .1, linetype = "solid", size = .5) +
  geom_errorbar(data = filter(exptdata_means_bycondition, condition == "Feature Search"), aes(x = dType, y = m, ymin = m-se, ymax = m+se), colour = "black", width = .15) +
  geom_line(data = filter(exptdata_means_bycondition, condition == "Feature Search"), aes(x = dType, y = m, group = 1), colour = "black") + 
  geom_point(data = filter(exptdata_means_bycondition, condition == "Feature Search"), aes(x = dType, y = m), colour = "black", size = 3) +
  geom_signif(comparisons = list(c("High", "Low"), c("Low", "Absent")), annotations = c("***","***"), y_position = c(.61,.57), colour = "black", tip_length = 0) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), expand = c(0,0), limits = c(0,.8)) +
  scale_x_discrete(labels = c("High\nreward", "Low\nreward", "Distractor\nabsent")) +
  scale_fill_manual(values = vmacCols) +
  scale_shape_manual(name = "Distractor Type", guide = guide_legend(), values = c(21,22,23)) +
  labs(y = "% Reward Omissions", x = "Trial Type", tag = "B", title = "Feature Search") +
  my_theme

omissionPlot_ms <- omissionPlot_singleton / omissionPlot_feature

save_plot(here::here("analysis", "figures", "omission_plot.png"), omissionPlot_ms, ncol = 1, nrow = 2, type = "cairo", base_aspect_ratio = 1.1, dpi = 300)

omissionPlot_ms

```

### Effect of Reward
Effect of reward ANOVA:
```{r Effect of Reward - Omissions}
omissions_vmac.ANOVA <- aov_car(m ~ condition*dType + Error(sub/dType), data = filter(exptdata_means_byparticipant, dType != "Absent"), anova_table = list(es="pes"))
```
Significant main effect of reward (indicating VMAC effect), but no interaction with or main effect of search condition.

Planned t-tests:
```{r Effect of Reward - Omissions - Planned t-tests}
singleton_high_omissions <- exptdata_means_byparticipant %>%
  ungroup() %>%
  filter(condition == "Singleton Search") %>%
  filter(dType == "High") %>%
  select(m) %>%
  pull()

singleton_low_omissions <-  exptdata_means_byparticipant %>%
  ungroup() %>%
  filter(condition == "Singleton Search") %>%
  filter(dType == "Low") %>%
  select(m) %>%
  pull()

feature_high_omissions <- exptdata_means_byparticipant %>%
  ungroup() %>%
  filter(condition == "Feature Search") %>%
  filter(dType == "High") %>%
  select(m) %>%
  pull()

feature_low_omissions <-  exptdata_means_byparticipant %>%
  ungroup() %>%
  filter(condition == "Feature Search") %>%
  filter(dType == "Low") %>%
  select(m) %>%
  pull()

singleton_VMAC_omissions.ttest <- t.test(singleton_high_omissions, singleton_low_omissions, paired = T)
singleton_VMAC_omissions.es <- dz_calculator(singleton_high_omissions, singleton_low_omissions)
feature_VMAC_omissions.ttest <- t.test(feature_high_omissions, feature_low_omissions, paired = T)
feature_VMAC_omissions.es <- dz_calculator(feature_high_omissions, feature_low_omissions)

vmac_singleton <- exptdata_means_byparticipant %>%
  filter(dType != "Absent", condition == "Singleton Search") %>%
  spread(dType, m) %>%
  mutate(vmac = High - Low) %>%
  pull(vmac)

vmac_feature <- exptdata_means_byparticipant %>%
  filter(dType != "Absent", condition == "Feature Search") %>%
  spread(dType, m) %>%
  mutate(vmac = High - Low) %>%
  pull

vmac_condition_bayes <- ttestBF(x = vmac_singleton, y = vmac_feature)
```
